---
title: "Religion-Voter Turnout Project"
output:
  pdf_document: default
  PDF: default
---

## Introduction


### Read in data
```{r}
library(readxl)
religion <- read_excel("religion_census.xlsx")
head(religion)
```
We only want data on predominantly African-American religious institutions to analyze the effects of Black Churches on voter turnout in the South. And we will narrow the data to a few states in the deep south: South Carolina, Georgia, Louisiana, Alabama, and Mississippi.

To understand what churches have the greatest impact, the Conference of National Black Churches provides a starting point. The organization said it represents more than 80% of African-American Christians, or around 20 million people in the U.S. It is made up of seven predominantly African-American Churches:

- African Methodist Episcopal Church (AME)
- African Methodist Episcopal Zion Church (AMEZ)
- Christian Methodist Episcopal Church (CME)
- Church of God In Christ (COGIC)
- National Baptist Convention of America, Inc., International (NBCA)
- National Baptist Convention, U. S. A. Inc. (NBC USA)
- Progressive National Baptist Convention, Inc. (PNBC) 

However, the convention does not include Black Protestants, which we will add on to the dataset.
We now go into the dataset to filter for these denominations.

### Filtering data
```{r}
library(dplyr)
library(tidyr)
#identify the abbreviations of 5 states we want to include
southern_states <- c("SC", "GA", "LA", "AL", "MS")
#identified the denominations' codes in the codebook
aa_denominations <- c("AMECNG", "AMEADH", "AMERATE", "AMEZCNG", "AMEZADH", "AMEZRATE", "CMECNG", "CMEADH", "CMERATE", "CGCCNG", "CGCADH", "CGCRATE", "NBCACNG", "NBCAADH", "NBCARATE", "NBCCNG", "NBCADH", "NBCRATE", "PNBCCNG", "PNBCADH", "PNBCRATE", "BPRTCNG", "BPRTADH", "BPRTRATE")
other_variables <- c("TOTCNG", "TOTADH", "TOTRATE", "FIPS", "STABBR", "CNTYNAME", "CNTYCODE", "POP2010")

southern_aa <- religion[c(other_variables, aa_denominations)] %>% 
  filter(STABBR  %in% southern_states) #data frame with all the African-American churches in the 5 states
```

#### Sanity check
```{r}
#we check if the number of counties in the dataframe matches up with the number of counties there are in the 5 states 
#LA has 64 states, AL has 67, GA has 159, SC has 46 and MS has 82.
nrow(southern_aa) == 64 + 67 + 159 + 46 + 82

#we also check whether total adherents are less than total adherents in the county
adh_var <- c("AMEADH", "AMEZADH", "CMEADH", "CGCADH", "NBCAADH", "NBCADH", "PNBCADH", "BPRTADH")
for (i in adh_var) {
  if (any(southern_aa[[i]] > southern_aa$TOTADH, na.rm = TRUE)) {i}
}
#check if total adherents are less than total population
which(southern_aa$TOTADH > southern_aa$POP2010)
```

### Clean up NAs
```{r}
library(visdat)
vis_miss(southern_aa) 
```

We observe that the dataset has a lot of missing data, attributed to the fact that some counties simply do not have an predominantly African-American church present.
*Note: There are a few denominations that has a higher missing data rate for their rate of adherents than they do for the total congregations variable. Namely: CGCC (Church of God in Christ), NBCA (National Baptist Convention of America Inc.) NBC (National Baptist Convention U.S.A), PNBC (Progressive National Baptist Convention, Inc.), BPRT (Black Protestants)*
We now examine what accounts for those inconsistencies.

```{r}
cng_var_miss <- c("CGCCNG", "NBCACNG", "NBCCNG", "PNBCCNG", "BPRTCNG")
adh_var_miss <- c("CGCADH", "NBCAADH", "NBCADH", "PNBCADH", "BPRTADH")
rate_var_miss <- c("CGCRATE", "NBCARATE", "NBCRATE", "PNBCRATE", "BPRTRATE")
check_rate <- vector(length = 3)
for (i in 1:5){
  print(southern_aa[!is.na(southern_aa[cng_var_miss[i]]) & is.na(southern_aa[rate_var_miss[i]]), c(cng_var_miss[i], adh_var_miss[i], rate_var_miss[i])])
}
#the adherents numbers in these counties' denominations are 0, therefore we can code the rate as 0.
```


```{r}
cng_var <- c("AMECNG", "AMEZCNG", "CMECNG", "CGCCNG", "NBCACNG", "NBCCNG", "PNBCCNG", "BPRTCNG")
table(apply(is.na(southern_aa[cng_var]), 1, sum))
```

The table examines how many denominations are absent in any given county. We find that 7 counties have all denominations present and 8 counties has none. 

```{r}
southern_aa[is.na(southern_aa)] <- 0
#verify that the dataset now has no missing values
```
We want to combine all the data on the denominations into one coherent set of data for the number of predominently African-American congregations, adherents and adherent rates for each county. This makes it easier to conduct analysis but also could expose the data to variations across different denominations. 

```{r}
rate_var <- c("AMERATE", "AMEZRATE", "CMERATE", "CGCRATE", "NBCARATE", "NBCRATE", "PNBCRATE", "BPRTRATE")
cng_var <- c("AMECNG", "AMEZCNG", "CMECNG", "CGCCNG", "NBCACNG", "NBCCNG", "PNBCCNG", "BPRTCNG")
southern_aa$AA_CNG <- apply(southern_aa[cng_var], 1, sum)
southern_aa$AA_ADH <- apply(southern_aa[adh_var], 1, sum)
southern_aa$AA_RATE <- apply(southern_aa[rate_var], 1, sum)
```

### EDA on the Religion dataset
```{r}
library(ggplot2)
#boxplots of denominations' adherents rate in counties
rate_var <- c("AMERATE", "AMEZRATE", "CMERATE", "CGCRATE", "NBCARATE", "NBCRATE", "PNBCRATE", "BPRTRATE")
rate_data <- southern_aa %>% 
  gather(key = "denomination", value = "rate", c(11, 14, 17, 20, 23, 26, 29, 32))
ggplot(data = rate_data, aes(x = denomination, y = rate)) + 
  geom_boxplot() + 
  labs(x = "Name of Denomination", y = "Rate of adherents per 1000 people")
  
ggplot(data = southern_aa, aes(x = STABBR, y = AA_RATE)) +
  geom_boxplot() + 
  labs(x = "State", y = "Rate of adherents per 1000 people") + 
  geom_hline(aes(yintercept = 50, color = "red"))
  
```
### Import election result data by county
```{r}
library(visdat)
pres_county <- read.csv("countypres_2000-2020.csv")
vis_miss(pres_county)
which(is.na(pres_county$county_fips)) %in% southern_aa$FIPS
#none of the missing fips are in the southern states dataset
pres_southern <- right_join(pres_county, southern_aa, by = c("county_fips" = "FIPS"))
pres_clean <- pres_southern %>% 
  filter(party == "DEMOCRAT") %>% 
  mutate(dem_share = candidatevotes/totalvotes) %>% 
  select(-c("office", "candidate", "party", "version", "mode", "county_name", "state_po", "state"))
```

### Visualize
```{r}
outcome <- pres_clean$dem_share > 0.5
outcome_char <- as.character(outcome)
outcome_char[outcome] <- "Majority"
outcome_char[!outcome] <- "Minority"
outcome_char <- factor(outcome_char, levels = c("Minority", "Majority"))
g_pres_church <- ggplot(data = pres_clean, aes(x = AA_RATE, y = dem_share))
g_pres_church + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = outcome_char, size = AA_ADH)) + 
  guides(alpha = "none") + 
  labs(x = "Number of Predominantly African-American Churches' Adherents per 1000 people in the county", y = "Share of votes for the Democratic candidate", color = "Democratic Vote",size = "Total Adherents")
```

#2008 & 2012 performance of Presidential candidates
```{r}
#2008 vote
pres_2008 <- pres_clean %>% 
  filter(year == 2008)

#encode the outcome variable as dem majority/minority
outcome_2008 <- pres_2008$dem_share > 0.5
outcome_2008_char <- as.character(outcome_2008)
outcome_2008_char[outcome_2008] <- "Majority"
outcome_2008_char[!outcome_2008] <- "Minority"
outcome_2008_char <- factor(outcome_2008_char, levels = c("Minority", "Majority"))

#graph rate vs. dem_share
g_pres_church_2008 <- ggplot(data = pres_2008, aes(x = AA_RATE, y = dem_share))
g_pres_church_visa2008 <- g_pres_church_2008 + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = outcome_2008_char, size = POP2010)) + 
  guides(alpha = "none") + 
  labs(x = "Number of Predominantly African-American Churches' Adherents per 1000 people in the county", y = "Share of votes for Obama", color = "Democratic Vote Share", size = "County Population", title = "2008 Presidential Election Vote by County")
#ggsave("scatterplot_2018.png", width = 10, height = 6)

#2012 vote
pres_2012 <- pres_clean %>% 
  filter(year == 2012)

#encode the outcome variable as dem majority/minority
outcome_2012 <- pres_2012$dem_share > 0.5
outcome_2012_char <- as.character(outcome_2012)
outcome_2012_char[outcome_2012] <- "Majority"
outcome_2012_char[!outcome_2012] <- "Minority"
outcome_2012_char <- factor(outcome_2012_char, levels = c("Minority", "Majority"))

#graph rate vs. dem_share
g_pres_church_2012 <- ggplot(data = pres_2012, aes(x = AA_RATE, y = dem_share))
g_pres_church_visa <- g_pres_church_2012 + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = outcome_2012_char, size = POP2010)) + 
  guides(alpha = "none") + 
  labs(x = "Number of Predominantly African-American Churches' Adherents per 1000 people in the county", y = "Share of votes for Obama", color = "Democratic Vote Share", size = "County Population", title = "2012 Presidential Election Vote by County")
g_pres_church_visa
#ggsave("scatterplot.png", width = 10, height = 6)
```

##### Try fitting the model
```{r}
fit <- lm(pres_2012$dem_share ~ pres_2012$AA_RATE)
summary(fit)
fit2 <- lm(pres_2008$dem_share ~ pres_2008$AA_RATE)
summary(fit2)
```

## Case Study: Louisiana
```{r}
Louisiana_2012_raw <- read_excel("LA_2012_1106_sta.xls", sheet = 1)
Louisiana_2012_raw <- Louisiana_2012_raw %>% 
  slice(-c(1:6, 9)) %>% 
  select(-c(1,3,5,7,9,11,13,15,17,19))

colnames(Louisiana_2012_raw) <- c("Jurisdiction", "Total", "Black_total_reg", "Dem_total_reg", "Black_dem_reg", "Rep_total_reg", "Black_rep_reg", "Third_total_reg" , "Black_third_reg") 

Louisiana_2012_raw1 <- Louisiana_2012_raw %>% slice(-c(1,2))

Louisiana_2012_voting_total <- Louisiana_2012_raw1 %>% 
  filter(Jurisdiction == "# VOTING") %>% 
  select(-1) %>% 
  apply(MARGIN = 2, FUN = as.integer) %>% 
  as.data.frame()
colnames(Louisiana_2012_voting_total) <- c("Total_vote", "Black_total_vote", "Dem_total_vote", "Black_dem_vote", "Rep_total_vote", "Black_rep_vote", "Third_total_vote", "Black_third_vote")

Louisiana_2012_vote_prop <- Louisiana_2012_raw1 %>% 
  filter(Jurisdiction == "% VOTING") %>% 
  select(-1) %>% 
  apply(MARGIN = 2, FUN = as.integer) %>% 
  as.data.frame()
colnames(Louisiana_2012_vote_prop) <- c("Total_vote_prop", "Black_total_vote_prop", "Dem_total_vote_prop", "Black_dem_vote_prop", "Rep_total_vote_prop", "Black_rep_vote_prop", "Third_total_vote_prop", "Black_third_vote_prop")

Louisiana_2012_registration <- Louisiana_2012_raw1 %>% 
  filter(Jurisdiction != "# VOTING" & Jurisdiction != "% VOTING")

Louisiana_2012_general <- cbind(Louisiana_2012_registration, Louisiana_2012_voting_total, Louisiana_2012_vote_prop)
Louisiana_2012_general$Total <- as.numeric(Louisiana_2012_general$Total)
```
### Plotting Louisiana dem_share against rate of predonimantly AA churches adherents
```{r}
pres_LA_2012 <- pres_clean %>% 
  filter(STABBR == "LA" & year == 2012) %>% 
  select(-c("CNTYNAME", "year")) %>% 
  mutate(Jurisdiction = Louisiana_2012_general[-1,]$Jurisdiction)

Louisiana_general_AA <- right_join(Louisiana_2012_general[-1,], pres_LA_2012, by = "Jurisdiction")

#graph rate vs. dem_share
outcome_LA_2012 <- Louisiana_general_AA$dem_share > 0.5
outcome_LA_2012_char <- as.character(outcome_LA_2012)
outcome_LA_2012_char[outcome_LA_2012] <- "Majority"
outcome_LA_2012_char[!outcome_LA_2012] <- "Minority"
outcome_LA_2012_char <- factor(outcome_LA_2012_char, levels = c("Minority", "Majority"))

g_pres_LA_2012 <- ggplot(data = Louisiana_general_AA, aes(x = AA_RATE, y = dem_share))
g_pres_LA_2012 + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = outcome_LA_2012_char, size = POP2010)) + 
  guides(alpha = "none") + 
  labs(x = "Number of Predominantly African-American Churches' Adherents per 1000 people in the county", y = "Share of votes for Obama", color = "Democratic Vote Share", size = "County Population", title = "2012 Presidential Election Vote in Louisiana by County")
```
### Plot Turnout against Rate of Adherents to Predominantly African-American Churches
```{r}
g_turnout_LA_2012 <- ggplot(data = Louisiana_general_AA, aes(x = AA_RATE, y = Black_total_vote_prop))
g_turnout_LA_2012 + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = dem_share, size = AA_ADH)) + 
  guides(alpha = "none") + 
    labs(x = "Rate of Adherents to Predominantly African-American Churches' Adherents (per 1000)", y = "Turnout to Registered Black voter Proportions (in %) ", color = "Democratic Vote Share", size = "Total Adherents", title = "2012 Presidential Election African-American Turnout in Louisiana by County") + 
  scale_color_gradient(low = "red", high = "blue")
ggsave("louisiana_turnout.png", width = 10, height = 6)
```
```{r}
g_turnout_LA_2012_dem <- ggplot(data = Louisiana_general_AA, aes(x = AA_RATE, y = Black_dem_vote_prop))
LA_turnout_2021_dem <- g_turnout_LA_2012_dem + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = dem_share, size = AA_ADH)) + 
  guides(alpha = "none") + 
  labs(x = "Rate of Adherents to Predominantly African-American Churches' Adherents (per 1000)", y = "Turnout among Registered Black voter Proportions (in %) ", color = "Democratic Vote Share", size = "Total Adherents", title = "Turnout among registered African-American Democrats in Louisiana's 2012 General Election by County") + 
  scale_color_gradient(low = "red", high = "blue")

g_turnout_LA_2012_rep <- ggplot(data = Louisiana_general_AA, aes(x = AA_RATE, y = Black_rep_vote_prop))
LA_turnout_2021_rep <- g_turnout_LA_2012_rep + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = dem_share, size = AA_ADH)) + 
  guides(alpha = "none") + 
  labs(x = "Rate of Adherents to Predominantly African-American Churches' Adherents (per 1000)", y = "Turnout among Registered Black voter Proportions (in %) ", color = "Democratic Vote Share", size = "Total Adherents", title = "Turnout among registered African-American Republicans in Louisiana's 2012 General Election by County") + 
  scale_color_gradient(low = "red", high = "blue")

g_turnout_LA_2012_third <- ggplot(data = Louisiana_general_AA, aes(x = AA_RATE, y = Black_third_vote_prop))
LA_turnout_2021_third <- g_turnout_LA_2012_third + 
  geom_smooth(method = "lm") + 
  geom_point(aes(alpha = 0.05, color = dem_share, size = AA_ADH)) + 
  guides(alpha = "none") + 
  labs(x = "Rate of Adherents to Predominantly African-American Churches' Adherents (per 1000)", y = "Turnout among Registered Black voter Proportions (in %) ", color = "Democratic Vote Share", size = "Total Adherents", title = "Turnout among registered African-American Third-party voters in Louisiana's 2012 General Election by County") + 
  scale_color_gradient(low = "red", high = "blue")
par(mfrow = c(1, 3))
LA_turnout_2021_dem
ggsave("louisiana_turnout_dem.png", width = 10, height = 6)
LA_turnout_2021_rep
ggsave("louisiana_turnout_rep.png", width = 10, height = 6)
LA_turnout_2021_third
ggsave("louisiana_turnout_third.png", width = 10, height = 6)
```



